Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX Format

### 2017-12-02 paper.tex

Future extreme heterogeneous systems will be managed by runtimes that handle the
complexities of abstracting across systems with diverse sets of compute
elements, deep memory hierarchies, and asymmetrical networks.  The increased
role of runtimes in resource and execution management is a direct result of the
complexity involved in designing applications that are able to portably extract
performance from complex architectures.  Often the target of domain-specific
languages, such runtimes exploit application-specific semantics and use the
state of the current system (e.g., hardware, software, configuration, etc.) to
optimize all aspects of distributed execution. Runtimes such as Legion [CITE]
expect to manage the entire compute infrastructure, but real-world constraints
require access to mass storage to be shared since it is used across applications
within organizations to relieve memory pressure, archive results, and provide
critical fault-tolerant capabilities. As a result, it is unlikely that in the
near-term the entire storage system hierarchy will be managed by a single
application runtime. Thus, the ability of application runtimes to optimize
execution is partly limited by the capabilities of I/O interfaces.

@@

Both HPC and cloud-based systems are seeing rapid advancements towards
heterogeneity. For instance, Amazon Web Services offers a large array of node
and network capabilities including FPGA and GPU co-processors. Emerging storage
technologies such as NVMe and persistent memories are offered by cloud providers
and will be integrated in future HPC systems, but even existing storage systems
are struggling to exploit their full performance [NOVA]. In fact, the increased
performance of next-generation media and networks suggests that future systems
may require a CPU per storage device [Slide Ref]. And in order to support a
dynamic range of interfaces, internal sub-systems will need to be generic and
re-usable to support the diverse of demands of runtime systems.

>>>>

Future extreme heterogeneous systems will be managed by runtimes that handle the
complexities of abstracting across systems with diverse sets of compute
elements, deep memory hierarchies, and asymmetrical networks.  The increased
role of runtimes in resource and execution management is a direct result of the
complexity involved in designing applications that are able to portably extract
performance from complex architectures.  Often the target of domain-specific
languages, such runtimes exploit application-specific semantics and use the
state of the current system (e.g., hardware, software, configuration, etc.) to
optimize all aspects of distributed execution. Runtimes such as Legion~\cite{bauer:sc12-legion}
expect to manage the entire compute infrastructure, but real-world constraints
require access to mass storage to be shared since it is used across applications
within organizations to relieve memory pressure, archive results, and provide
critical fault-tolerant capabilities. As a result, it is unlikely that in the
near-term the entire storage system hierarchy will be managed by a single
application runtime. Thus, the ability of application runtimes to optimize
execution is partly limited by the capabilities of I/O interfaces.

@@

Both HPC and cloud-based systems are seeing rapid advancements towards
heterogeneity. For instance, Amazon Web Services offers a large array of node
and network capabilities including FPGA and GPU co-processors. Emerging storage
technologies such as NVMe and persistent memories are offered by cloud providers
and will be integrated in future HPC systems, but even existing storage systems
are struggling to exploit their full performance~\cite{xu:fast16-nova}. In fact, the increased
performance of next-generation media and networks suggests that future systems
may require a CPU per storage device~\cite{samuels:oss16}. And in order to support a
dynamic range of interfaces, internal sub-systems will need to be generic and
re-usable to support the diverse of demands of runtime systems.
